---
title: "Developer-instructions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developer-instructions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wqbench)
```

## Updating the Add Data Template

The instructions in the uploading data template need to up updated when any changes to the reference data sheets occur. 

- trophic-groups.csv
  - need to update trophic_group column and ecological_group column instructions if new groups are added or groups are removed
  - the groups have been manually added to the instruction tab
  - the validation is done off of the trophic-groups.csv so it is possible for the instructions to get out of sync with the allowed values
   
- lifestage-codes.csv
  - need to update simple_lifestage column instructions if new simple lifestages are added or removed 
  - the validation is done off of the lifestage-codes.csv so it is possible for the instructions to get out of sync with the allowed values
  
- concentration-endpoints.csv
  - need to update endpoint column instructions if endpoints are added or removed
  - the list of endpoints on the third sheet are generated using the function XXX, this needs to be run each time the concentration-endpoints.csv file is updated
  - the validation is done off of the concentration-endpoints.csv so it is possible for the instructions to get out of sync with the allowed values  
  
To regenerate the list of endpoints you will need to go to the *data-raw/template.R* file and run the script.
This will add the template section to the package and regenerate the list of allowed endpoint codes.

In the templates folder only the csv files should be updated manually.
The excel file is for people to use when filling in with their data and is generated from running the *data-raw/template.R* script.