---
title: "Developer Instructions"
output:
  rmarkdown::github_document:
  keep_md: true
vignette: >
  %\VignetteIndexEntry{Developer Instructions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<!--
rmarkdown::html_vignette:
rmarkdown::github_document:

rmarkdown::render("vignettes/Developer-instructions.Rmd")
-->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Updating Reference Data for the Database

Each time the Ecotox database is updated, the reference files in the package
should be reviewed for necessary changes. This process involves generating
some files to illustrate changes, having those files reviewed by
an expert, and updated if necessary, and then updating the data
files in the package with the new changes.

These changes may involve:

- concentration conversions and duration conversions for newly-added
  concentration and duration units,
- trophic groups and life stage codes for newly added species.

Lists of BC species and concentration endpoints are unlikely to require changes
but should be reviewed periodically.

### Process

0. Download (clone or pull) the `wqbench` repo and open in RStudio. The best practice
  is to open a new git branch, called something like `update-[date-of-new-ecotox]`.
  In Rstudio, git pane -> New branch -> give it a name, and check "Sync branch
  with remote".

1. Ensure you have the most recent copy of the database with the reference files
added. To do this, load the package, and run the function `wqb_create_data_set()`:

```{r, eval=FALSE}
devtools::load_all()

wqb_create_data_set(
  file_path = "~/Ecotoxicology/ecotox",
  version = 1,
  folder_path = "~/Ecotoxicology/ecotox_db/"
)
```

2. Execute the code in the script *scripts/review-reference-datasets-01.R*.
  - This script will generate and save a set of csv files that need to be
    reviewed and updated to allow new values through the data cleaning steps.
  - At the top of the script you will need to set the file path for the database
    and the location to save the files that are generated. By defeault they
    will save at: `"~/Poisson/Data/wqbench/[current-year]/review/to-be-reviewed/"`

3. After the files are generated, they need to be [reviewed and updated](#review-process) by a
  technical expert.
  - The next step must wait until the review is complete.
  - The life stage code file can't be reviewed until after the trophic groups
    have been updated.

4. Once the files are reviewed they should be placed in:
  `"~/Poisson/Data/wqbench/[current-year]/review/completed/"`
  - Then run the *scripts/update-reference-datasets-01.R* script.
  - Run this line by line - You will be shown several displays showing the
    updates that will be made to the internal data - ensure these look ok. If
    they do, continue executing the script, if not you will need to revisit your
    review and the changes you made to the csv files in the `completed` folder.
  - This will read in the reviewed files and update the reference files in the
    *inst/extdata* folder.

5. The package needs to be re-built for the files to be part of the package.
  Run `devtools::load_all()`.

6. Run the `wqb_create_data_set()` function to create the database with the new
  reference data.

7. Repeat steps 2 through 6 but run *scripts/review-reference-datasets-02.R*,
  review the file, and run *scripts/update-reference-datasets-02.R*.
  - The life stage codes have to be generated seperately as they are based on
    the trophic groups. The trophic groups need to be updated before life stage
    codes can be reviewed.

8. Run the `wqb_create_data_set()` function again to create the database with
  the new reference data.

9. These steps will have caused changes to several files in `inst/extdata`. The
  changes will be shown in the Git pane in RStudio. Commit these files to Git,
  and push it to GitHub. Go to the GitHub repository, and open
  a pull request from your branch (created in step 1). Best practice is to have
  someone review the PR, but if you know the changes are good, you can merge it
  yourself. At the very least, ensure that all of the automated checks pass
  before you merge.

10. Once your update is merged, install the new version:
  `devtools::install_github("bcgov/wqbench")`.

### Review process

Below are instructions for how to fill out and complete each of the reference
files.

#### Concentration Conversion

This data set should be reviewed each time a new version of the ECOTOX database
is released.

- If there is no value in the **conc_conversion_flag** column this indicates it
  is a new concentration unit that was not in the previous version of the
  database.
- The goal of the review is to ensure all cells in the **conc_conversion_flag**
  column are filled in.
  - A zero (`0`) indicates the concentration cannot be converted to mg/L or ppm.
    - If the units cannot be converted to mg/L or ppm then put a `0` in the row.
    - No additional columns need to be filled in if the unit is given a value of
     `0` in the **conc_conversion_flag** column.
  - A one (`1`) indicates the concentration can be converted to mg/L or ppm.
    - If the units can be converted put a `1` in the row.
      - In the **conc_conversion_value_multiplier** column fill in the value
        needed to convert the units into mg/L or ppm.
      - In the **conc_conversion_unit** column fill in either `mg/L` or `ppm`.
  - Once completed, this file should be saved in the `"completed"` subfolder
      in the review folder.

If any incorrect conversions are found, then those rows can be updated.

#### Duration Conversion

This data set should be reviewed each time a new version of the ECOTOX database
is released.

- If there is no value in the **duration_units_to_keep** column this indicates
  it is a new concentration unit that was not in the previous version of the
  database.
- The goal of the review is to ensure all cells in the
  **duration_units_to_keep** column are filled in.
  - A zero (`0`) indicates the duration cannot be converted to hours.
    - If the units cannot be converted to hours, then put a `0` in the row.
    - No additional columns need to be filled in if the unit is given a value of
      `0` in the **duration_units_to_keep** column.
  - A one (`1`) indicates the duration can be converted to hours.
    - If the units can be converted put a `1` in the row.
      - In the **duration_value_multiplier_to_hours** column fill in the value
        needed to convert the units into hours.
      - In the **comments** column write why the conversion was chosen since
        there may be an assumption made during the conversion. For example, if
        converting month into hours are you basing the conversion on 30 or 31
        days.
  - Once completed, this file should be saved in the `"completed"` subfolder
      in the review folder.

If any incorrect conversions are found, then those rows can be updated.

#### Trophic Groups

This data set should be reviewed each time a new version of the ECOTOX database
is released.

- Two files will be generated to help review the trophic group data.
  - *missing-trophic-group-review.csv*
    - This is a summary that shows the unique phylum, class, order, and
      family of new taxa in the new version of Ecotox database, that do not meet
      the exclusion criteria. In other words, these are candidate taxa to be
      added to the internal dataset.
    - This file contains three columns for the reviewer to consider:
      **trophic_group**, **ecological_group**, and **exclude_from_db**.
        - If the taxon is not appropriate for inclusion, put a `1` in the
          **exclude_from_db** column. Otherwise leave it blank.
        - If it is appropriate to include, fill out the **trophic_group** and
          **ecological_group** columns. Valid values for **trophic_group** are:
          "Invertebrate", "Algae", "Amphibian", "Plant", "Bacteria", "Fish".
          Valid values for **ecological_group** are: "Planktonic Invertebrate",
          "Other", "Salmonid".
        - If there is a row in the "missing-trophic-review" sheet that doesn't
          have family or order, do one of two things: if the phylum/division is
          really small and/or the whole taxon can be assigned an ecological group
          and trophic group, do so. If not (e.g. Annelida, Arthropoda), put `1`
          in the **exclude_from_db** column so that high-level taxon won't appear
          for you to review again, but lower levels within that phylum/division
          will. It is also likely that lower taxonomic entities have
          been assigned to the appropriate trophic group and ecological
          group.
    - Once completed, this file should be saved in the `"completed"` subfolder
      in the review folder.
  - *species-coded-in-db-ref.csv*
    - This is a list of all the species data from the database that have been
      filtered where **organism_habitat** is "Water".
    - This file is to help find which **class** and **tax_order** (order) do not
      have coding for the ecological or trophic groups.
    - This file is for reference and not to be updated or sent back for
      integration.

#### Life Stage Codes

This data set should be reviewed each time a new version of the ECOTOX database
is released.

This data set will be sent separately after the first round of files is reviewed
because this data depends on the updates to the trophic group data. Only fish
and amphibian groups need the life stage categorized into simple groups, so this
file only contains fish and amphibians.

*lifestage-code-review.csv*:
- If there is no value in the **simple_lifestage** column this indicates it is a
  new life stage that was not in the previous coding of the database.
- The goal of the review is to ensure all cells in the **simple_lifestage**
  are filled in.
  - In the **simple_lifestage** column fill in any empty cells with either `els`
    (early life stage), `juvenile` or `adult`.

Once completed, this file should be saved in the `"completed"` subfolder in the
review folder.

#### BC Species

This reference dataset has not been included in the *review-reference-datasets.R* script as it should not vary from year to year.

This data set is a comprehensive list that was generated from B.C. Conservation Data Centre.

Any species that is not listed in the *bc-species.csv* reference file is marked as not present in BC and thus it is robust to new species added to the ECOTOX dataset.

#### Concentration Endpoints

This reference dataset has not been included in the *review-reference-datasets.R* script as it should not vary from year to year.

There is a comprehensive list that should not need to be adjusted.

The list of concentration endpoints are generated in the *scripts/concentration-endpoints.R* script.

If updates are required then update the *concentration-endpoints.R* script.

### Updating the Add Data Template

**This will rarely need to be done as additions / removals to ecological groups
and trophic groups are uncommon**

#### Associated Reference Data

The instructions in the upload data template need to be updated when changes to the reference data sheets occur.

- *trophic-groups.csv*
  - Update **trophic_group** and **ecological_group** column instructions if new groups are added or groups are removed.
    - ie if `Plant` is removed as a trophic group or `Salmonid` is removed as an ecological group.
    - Since the groups (Plant, Bacteria, Amphibian, etc) have been written into the instruction tab this is why updates to the instructions are needed if new categories are added.
  - The validation of the values is done from the *trophic-groups.csv* which means it is possible for the instructions to get out of sync with the allowed values.
  - Adding new species to the *trophic-groups.csv* does not trigger a change to the instructions.

- *lifestage-codes.csv*
  - Update **simple_lifestage** column instructions if new simple life stages categories are added or removed.
    - ie if `els` is removed as a simple life stage.
  - The validation is done off the *lifestage-codes.csv* which means it is possible for the instructions to get out of sync with the allowed values.

- *concentration-endpoints.csv*
  - Update **endpoint** column instructions if endpoints are added or removed.
  - The list of endpoints are generated by the script *scripts/concentration-endpoints.R*.
  - The validation is done off of the *concentration-endpoints.csv* so it is possible for the instructions to get out of sync with the allowed values.

#### To update the template

1. Make any required changes to the trophic groups, life stage codes or concentration endpoints reference data.
2. Edit the *inst/template/template-data-data.csv* and *inst/template/template-data-instructions.csv* files.
   - Do not edit the file *inst/template/template-data.xlsx*.
3. To regenerate the template run the script *data-raw/template.R*.
4. Review the *inst/template/template-data.xlsx* to confirm the edits came through and the template looks as expected.
   - Do not edit the file *inst/template/template-data.xlsx*.

## Push and install the pacakge

Finally, after all of the all of the updates are completed, push them
to GitHub. Install the updated package with
`devtools::install_github("bcgov/wqbench")`.
