---
title: "Developer-instructions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developer-instructions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wqbench)
```

## Updating the Internal Data Set in the Shiny app

The shiny app is located in the [shinywqbench](https://github.com/bcgov/shinywqbench) repository. 
To update the internal data file

1. Go to the inst/extdata/data.R file.
2. Run the script.

This needs to occur each time the ECOTOX EPA data set is updated on the website.  

## How to Deploy the Shiny App

The shiny app is located in the [shinywqbench](https://github.com/bcgov/shinywqbench) repository.

1. Go to the scripts/deploy.R file.
2. Run the script.

It is advised to first run the code that deploys the app name of `shinywqbench-dev` and confirm the app deploys and functions as expected.
This app is referred to as the development app. 

Once it has been confirmed that the app functions and deploys properly to then run the second chunk of code with the app name `shinywqbench`.
This app is referred to as the production app. 

This will help to ensure the production app is always in a working state.

## Updating the Add Data Template

The instructions in the uploading data template need to up updated when any changes to the reference data sheets occur. 

- trophic-groups.csv
  - need to update trophic_group column and ecological_group column instructions if new groups are added or groups are removed
  - the groups have been manually added to the instruction tab
  - the validation is done off of the trophic-groups.csv so it is possible for the instructions to get out of sync with the allowed values
   
- lifestage-codes.csv
  - need to update simple_lifestage column instructions if new simple lifestages are added or removed 
  - the validation is done off of the lifestage-codes.csv so it is possible for the instructions to get out of sync with the allowed values
  
- concentration-endpoints.csv
  - need to update endpoint column instructions if endpoints are added or removed
  - the list of endpoints on the third sheet are generated using the function XXX, this needs to be run each time the concentration-endpoints.csv file is updated
  - the validation is done off of the concentration-endpoints.csv so it is possible for the instructions to get out of sync with the allowed values  
  
To regenerate the list of endpoints you will need to go to the *data-raw/template.R* file and run the script.
This will add the template section to the package and regenerate the list of allowed endpoint codes.

In the templates folder only the csv files should be updated manually.
The excel file is for people to use when filling in with their data and is generated from running the *data-raw/template.R* script.

## Updating Reference Data

### Download Most Recent Database

Start by downloading the most recent copy of the database and adding the current reference data to it

```{r, eval=FALSE}
wqb_create_data_set(
  file_path = "~/Ecotoxicology/ecotox",
  version = 1,
  folder_path = "~/Ecotoxicology/ecotox_db/"
)
```

### BC Species

This list should not vary or need to be reviewed often as it is not expected that the species in BC will vary from year to year.
This is a comprehensive list that was generated from B.C. Conservation Data Centre.

Any species that is not listed in the *bc-species.csv* reference file is marked as not present.
Thus this list is robust to new species added to the ECOTOX dataset. 

### Concentration Conversion

This data set should be reviewed each time a new pull from the ECOTOX dataset is generated. 

After the database is created. 

Run the script *scripts/review-concentration-conversion.R* to generate one file for review.

The script will generate one file:

1. Concentration conversion table from the database with the coding of whether the unit is retained and converted.
   - Values with 1 in the conc_conversion_flag column are retained and converted based on the value in the conc_conversion_value_multiplier column to the unit listed in the conc_conversion_unit column.
   - Look for values that have nothing in the conc_conversion_flag column.
   - To add new values that need to be added in the compiled dataset write the conversion value in the add_new_multipler column and the corresponding converted unit in the add_new_unit column.

### Concentration Endpoints

The list of concentration endpoints are generated in the *data-raw/concentration-endpoints.R* script. 

This is a comprehensive list that should not need to be adjusted unless a different set of endpoint codes allowed. 

### Duration Conversion

This data set should be reviewed each time a new pull from the ECOTOX data set is generated. 

After the database is created. 

Run the script *scripts/review-duration-conversion.R* to generate one file for review.

The script will generate one file:

1. Duration conversion table from the database with the coding of whether the unit is retained and converted.
   - If a 0 is present in the duration_unit column then those values are not retained in the final compiled dataset. Values with 1 are retained and converted based on the value in the duration_value_multiplier_to_hours column.
   - Look for values that have nothing in the duration_unit column.
   - To add new values that need to be added in the compiled dataset write the conversion value in the add_new_multipler column.
   - Add in the rational/assumption for the duration conversion in the add_comments column.
   
### Lifestage Codes

This data set should be reviewed each time a new pull from the ECOTOX data set is generated. 

After the database is created. 

Run the script *scripts/review-lifestage-codes.R* to generate one file for review.

The script will generate one file:

1. Life stage codes from the database with their coding to the simple life stages. 
   -  For values that need to be added to the reference dataset, write the simple life stage in the add column.

### Trophic Groups

This data set should be reviewed each time a new pull from the ECOTOX data set is generated. 

After the database is created.

Run the script *scripts/review-trophic-groups.R* to generate two files for review.

The script will generate three files:

1. Species data from the database. 
   - Extra addition should be provided to species marked with a 1 in the be species column as these are species present in BC that are categorized into trophic groups.
   - This file is provided for reference but should not be used for listing the new groups to add to the reference dataset. 
2. A condensed list of the phylum, class, order and family of all the species that were not categorized with a trophic group.
  -  For groups that need to be added to the reference data set these need to `TRUE` written in the add column. The class and tax_order column will be joined to the class and order column in the reference dataset, they need to match identically or they can not be joined. 
  - This file will be used to add the new trophic groups to the current reference file. 
3. The current reference data joined to the filtered and cleaned class and order list in the database. 

These files need to be reviewed and will guide any updates to the reference data for the trophic groups.
Update the file *inst/extdata/trophic-groups.csv* based on the review.

### After All Reference Datasets are Review

Once all the reference data sets have been reviewed and updated (as needed), re-build the package. 

Then restart your R session and create the dataset again. 

```{r, eval=FALSE}
wqb_create_data_set(
  file_path = "~/Ecotoxicology/ecotox",
  version = 1,
  folder_path = "~/Ecotoxicology/ecotox_db/"
)
```
